---
title: "Project 2"
author: "Group 5: Katelyn Wnuk, Brennan Chan, Emmett Collins"
format: pdf
editor: visual
---
Load in the data keeping only numeric values for an easier pipeline.
```{r, message=FALSE, warning=FALSE}
library(tidymodels)
library(tidyverse)
library(car)
library(olsrr)
library(dplyr)
library(GGally)

data_car = read_csv("http://www.jpstats.org/data/car_price.csv") |>
  select( where(~ !is.character(.x)) )
```

## REWRITE ME 
There are a several methods for picking what variables we want to generate a linear model with. One of these methods is to look at all possible models and pick the best one based on a certain criteria. In this case we will use adjusted R-squared to determine the best models.
```{r}
fit = lm(price ~ ., data = data_car)
subsets_best = ols_step_best_subset(fit) 
plot(subsets_best)
subsets_best
```
We decided to pick the model with 14 predictor variables (car_ID, symboling, wheelbase, carlength, carwidth, carheight, curbweight, enginesize, stroke, compressionratio, horsepower, peakrpm, citympg, highwaympg )




# temp BRENNAN CHAN DELETE ME 
Changed how we view the best subsets for better viewing and more closely resembling his notes.