---
title: "Homework 4"
author: "Group 1: Brennan Chan, Connor Alfeiri, Tyler Frankowiak"
format: pdf
editor: visual
---

# Homework 4

## Question 1

Although invisible and intangible, ambient conditions such as air quality ($x_1$), temperature ($x_2$), odor/aroma ($x_3$), music ($x_4$), noise level ($x_5$), and overall image ($x_6$) may affect guests’ satisfaction with their stay at a hotel. A study in the *Journal of Hospitality Marketing & Management* (Vol. 24, 2015) was designed to assess the effect of each of these ambient factors on customer satisfaction with the hotel ($y$). Using a survey, researchers collected data for a sample of 422 guests at 5-star hotels. All variables were measured as an average of several 5-point questionnaire responses. The results of the multiple regression are summarized in the accompanying table.

| **Variable**      | $\beta$ estimate |
|-------------------|------------------|
| Air Quality       | 0.122            |
| Temperature       | 0.018            |
| Odor/Aroma        | 0.124            |
| Music             | 0.119            |
| Noise/Sound Level | 0.101            |
| Overall Image     | 0.463            |

### Part A

Write the equation of a first-order model for hotel image ($y$) as a function of the six ambient conditions.

$$
\hat{y} = (\text{unknown intercept})\beta_0 +
  0.122\,(x_1) +
  0.018\,(x_2) + \\
  0.124\,(x_3) + 
  0.119\,(x_4) +
  0.101\,(x_5) +
  0.463\,(x_6) +
  \varepsilon 
$$

### Part B

Give a practical interpretation of each of the ($\beta$)-estimates shown.

Each ($\beta$)-estimate represents the *expected change in the predicted hotel image rating* (($\hat{y}$)) for a **one-unit increase** in the corresponding ambient condition, **holding all other variables constant**.

-   ($\beta_{1}$ = 0.122): For every one-unit increase in **Air Quality**, the predicted hotel image ($\hat{y}$) increases by 0.122, assuming other factors remain constant.\
-   ($\beta_{2}$ = 0.018): For every one-unit increase in **Temperature**, ($\hat{y}$) increases by 0.018, holding all else constant.\
-   ($\beta_{3}$ = 0.124): For every one-unit increase in **Odor/Aroma**, ($\hat{y}$) increases by 0.124, holding all else constant.\
-   ($\beta_{4}$ = 0.119): For every one-unit increase in **Music**, ($\hat{y}$) increases by 0.119, holding all else constant.\
-   ($\beta_{5}$ = 0.101): For every one-unit increase in **Noise/Sound Level**, ($\hat{y}$) increases by 0.101, holding all else constant.\
-   ($\beta_{6}$ = 0.463): The intercept represents the predicted value of ($\hat{y}$) (hotel image) when all six ambient condition variables are equal to zero.

## Question 2

Consider a multiple-regression model for predicting the total number of runs scored by a Major League Baseball (MLB) team during a season. The data on number of walks ($x_1$), singles ($x_2$), doubles ($x_3$), triples ($x_4$), home runs ($x_5$), stolen bases ($x_6$), times caught stealing ($x_7$), strike-outs ($x_8$), and ground outs ($x_9$) and total number of runs scored ($y$) for each of the 30 teams during the 2017 MLB season can be read from [Link](http://www.jpstats.org/Regression/data/MLBRUNS2017.csv)

### Part A

Fit the multiple regression model to the data. Give the fitted model equation.

```{r, Messages=FALSE}
# Messages = False sometimes doesn't work so I added insurance
suppressPackageStartupMessages(library(tidymodels))
suppressPackageStartupMessages(library(tidyverse))

# quick import of the data set
# got some warning when trying to read in the csv so suppressed it with the show col types 
dat = read_csv("http://www.jpstats.org/Regression/data/MLBRUNS2017.csv", show_col_types = FALSE)

# Get only the usefull columns 
mlb_data <- dat[, c("RUNS", "WALKS", "SINGLES", "DOUBLES", "TRIPLES", 
                    "HOMERUNS", "STOLEBASES", "CAUGHTSTEAL", "STRIKEOUTS", "GRNDOUTS")]
```

```{r}

# the '.' selects all but the previously listed column from a data variable
data_recipe = recipe(RUNS~., data = mlb_data)

# setup the model
lm_spec = linear_reg(engine = "lm")

# Create workflow object
wf = workflow()
wf = add_recipe(wf, data_recipe)
wf = add_model(wf, lm_spec)

# Fit the model to the data 
fit_lm = fit(wf, data = mlb_data)

# Get coefficients neatly
tidy_output = tidy(fit_lm)
print(tidy_output)

# might want to get rid of tidy output to "reduce duplicate code" 
```

Fitted Model Equation:

$$
\hat{y} = -709.082
+ 0.397\,(x_1)
+ 0.618\,(x_2)
+ 0.819\,(x_3)
+ 0.756\,(x_4)
+ 1.561\,(x_5) \\
+ 0.089\,(x_6)
+ 1.080\,(x_7)
+ 0.018\,(x_8)
+ 0.042\,(x_9)
$$

### Part B

Give practical interpretations of the estimated coefficients.

-   **WALKS (BB)**: For each additional walk, predicted runs increase by $0.3968$ runs on average ($p = 1.17\times 10^{-4}$, significant).

-   **SINGLES**: For each additional single, predicted runs increase by $0.6177$ runs on average ($p = 5.23\times 10^{-6}$, significant).

-   **DOUBLES**: For each additional double, predicted runs increase by $0.8192$ runs on average ($p = 2.47\times 10^{-4}$, significant).

-   **TRIPLES**: For each additional triple, predicted runs increase by $0.7557$ runs on average ($p = 0.172$, not statistically significant at 0.05).

-   **HOMERUNS**: For each additional home run, predicted runs increase by $1.5611$ runs on average ($p = 6.78\times 10^{-7}$, significant).

-   **STOLEBASES**: For each additional stolen base, predicted runs increase by $0.0892$ runs on average ($p = 0.722$, not significant).

-   **CAUGHTSTEAL**: For each additional caught stealing, predicted runs increase by $1.0796$ runs on average ($p = 0.218$, not significant).\
    *Note:* The positive sign here can occur due to correlation structure with other variables; interpret cautiously unless significant.

-   **STRIKEOUTS**: For each additional strikeout, predicted runs increase by $0.0175$ runs on average ($p = 0.736$, not significant).\
    *likely reflecting multicollinearity with other hitting stats.*

-   **GRNDOUTS/Ground Outs**: For each additional ground out, predicted runs increase by $0.0422$ runs on average ($p = 0.512$, not significant).

### Part C

Test for overall model utility.

```{r}
# Get model summary stats
glance_output = glance(fit_lm)
print(glance_output)
```

The p-value of the model is $1.668341 \times 10^{-9}$ which is less than $0.05$ the regression model provides evidence that at least one offensive variable contributes meaningfully to predicting total runs.
