---
title: "Homework2"
subtitle: "Group 10"
author: "Brennan Chan, INSERT YOUR NAME HERE"
format: pdf
editor: visual
---

# Problem 4

-   A contract engineer at DuPont Corp. studied the rate at which a spilled volatile liquid will spread across a surface (Chemical Engineering Progress, January 2005). Assume 50 gallons of methanol spills onto a level surface outdoors. The engineer used derived empirical formulas (assuming a state of turbulent free convection) to calculate the mass (y, in pounds) of the spill after a period of time (x) ranging from 0 to 60 minutes. The calculated mass values can be found at [link](http://www.jpstats.org/data/LIQUIDSPILL.csv).

```{r, message=FALSE}
library(tidyverse)
# changed to read directly from the link 
df_dupont = read.csv("http://www.jpstats.org/data/LIQUIDSPILL.csv")

# double check whats in here
head(df_dupont)
```

## A. Find the least squares line for the data.

```{r}
# Predicting the Mass using time
fit = lm(MASS~TIME,data=df_dupont)
summary(fit)
```

The fitted least squares regression line is: $$
E(y) = 5.2207 - 0.1140 \times TIME
$$

## B. Plot the data using ggplot and add the least squares line to the plot.

```{r}
df_dupont |> ggplot(aes(x = TIME, y = MASS))+
  geom_point() + 
  geom_smooth(method = "lm")
```

## C. Determine the utility of the model by testing the slope.

```{r}
summary(fit)
```

Since the p-value is very small, we reject H~0~ and conclude that TIME is a significant predictor of MASS. Thus, the regression model is useful for describing how the mass of the spill decreases over time.

## D. Comment on the fit of the model.

-   The regression model indicates a strong negative linear relationship between time and mass. The estimated slope is -0.114 (p \< 0.001) showing that mass decreases significantly as time increases. The model explains approximately 85% of the variation in mass (R² = 0.85), suggesting a good fit. The residual standard error of 0.86 is relatively small compared to the response scale, indicating that the model predicts well.

## E. Find a 90% confidence interval for the mean mass of all spills with an elapsed time of 15 minutes.

```{r}
# Confidence interval at TIME = 15
predict(fit, newdata = data.frame(TIME = 15), interval = "confidence", level = 0.90)
```

The 90% regression interval was $$[3.173,\; 3.848]$$

## F. Find a 90% prediction interval for the mass of a single spill with an elapsed time of 15 minutes. result.

```{r}
# Prediction interval at TIME = 15
predict(fit, newdata = data.frame(TIME = 15), interval = "prediction", level = 0.90)
```

The 90% prediction interval was $$[1.997,\; 5.024]$$
