---
title: "Regression Analysis Project 1"
author: "Brennan Chan, Tim McKinley, Connor Kelly"
format: pdf
editor: visual
---
# Project 1

# Fit a multiple regression model using Earnings as the dependent variable, and Events and Score as the independent variables

First lets load in the packages and data.

```{r, Messages=FALSE}
# Core packages (tidymodels but no pipes)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(tidymodels))

# Read the data (put Golf.csv in your working directory)
golf <- read.csv("Golf.csv", header = TRUE, stringsAsFactors = FALSE)

# Peek
str(golf)
summary(golf)

```

Now we can fit the multiple regression model

```{r}
# Model spec (no %>%): use set_engine(linear_reg(), "lm")
mod_spec <- set_engine(linear_reg(), "lm")

# Fit model
fit_multi <- fit(mod_spec, Earnings ~ Events + Score, data = golf)

# Pull the underlying lm object for classical outputs
lm_multi <- fit_multi$fit

# Coefficient table (t-tests) & overall fit (R^2, adj R^2, F-test)
coef_table <- broom::tidy(lm_multi)        # term, estimate, std.error, statistic (t), p.value
fit_glance <- broom::glance(lm_multi)      # r.squared, adj.r.squared, sigma, statistic (F), p.value, AIC, BIC

# ANOVA table (global F test details)
anova_multi <- anova(lm_multi)

coef_table
fit_glance
anova_multi

```

# Give the equation of the model and the equation of the fitted model for the fit in item 2

Model Form: E(Earnings∣Events,Score)= $\beta_0 + \beta_1 Events + \beta_2 Score$

Where:

-   $\beta_0$ = Intercept = 73,365,196

-   $\beta_1$ = Coefficient for Events = -116,049

-   $\beta_2$ = Coefficient for Score = -965,039

Fitted model: E(Earnings∣Events,Score)= $73,365,196−116,049(Events)−965,039(Score)$


